Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-08-29 main.tex
Update on Overleaf.

\begin{lemma}\label{lemma:boundarysimplices}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$. 
    \begin{itemize}
        \item Any face $F \in \Faces(\calT)$ is not contained in the boundary if and only if it is contained in exactly two $n$-simplices.
        \item Any face $F \in \Faces(\calT)$ is contained in the boundary if and only if it is contained in exactly one $n$-simplex.
        \item The simplices contained in the boundary constitute a triangulation of the boundary.
        % \item $\calT$ is non-branching, that is, every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$. 
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove these statements in several steps.
    \begin{enumerate}
    \item 
    We $\mathring \calM := \calM \setminus \partial \calM$ denote the interior of the manifold. 
    We will use the following fact: if $S \in \calT$ has an inner point that lies on $\partial \calM$, then all inner points of $S$ are on $\partial \calM$. 
    Since the boundary $\partial \calM$ is closed, every $S \in \calT$ is either a subset of the boundary or all its inner points lie in the interior $\mathring \calM$ of the manifold.

    \item 
    We recall an auxiliary result.
    Suppose that $Y$ is a topological space homeomorphic to a sphere of dimension $m$ and that $X \subseteq Y$ is homeomorphic to a sphere of dimension $m-1$, where $m \geq 1$. 
    As a consequence of the Jordan-Brouwer separation theorem~\cite[Corollary 5.24]{mayer1989algebraische}~\cite[Corollary 6.4]{massey1981algebraic}, 
    we know that $Y \setminus X$ has got two connected components.

    \item 
    We prove an auxiliary result. 
    Suppose that $Y$ is a topological space homeomorphic to a sphere of dimension $m$ and that $X \subseteq Y$ is homeomorphic to a sphere of dimension $m-1$, where $m \geq 1$. 
    The embedding of $X$ into $Y$ cannot be onto~\cite[Theorem~2.55]{lee2011topological}, and so there exists $y \in Y \setminus X$. 
    Using the homeomorphism from $Y$ into the $m$-sphere, which maps $y$ to the northpole without loss of generality, and a stereographic projection, 
    we see that $Y \setminus \{y\}$ is homeomorphic to $R^{m}$.
    As a consequence of the Jordan-Brouwer separation theorem~\cite[Theorem 5.23]{mayer1989algebraische}~\cite[Corollary 5.24]{mayer1989algebraische}~\cite[Corollary 6.4]{massey1981algebraic}, $Y \setminus (X \cup \{y\})$ has got two connected components,
    and thus $Y \setminus X$ has got two connected components.

    \todo{Decide which paragraph remains...}

@@

    Consider the case that $z_F$ lies on the boundary of $\calM$
    and suppose that $F$ is contained in $K$ distinct $n$-simplices of $\calT$.
    By adding at least one dimension, we can double the manifold $\calM$ along the boundary and obtain the doubled manifold $\calM'$. Similarly, we can construct a doubling of the triangulation $\calT'$ such that $F$ is contained in exactly $2K$ distinct $n$-simplices of $\calT'$. We know that $\calM'$ is a manifold without boundary, and hence $F$ is an interior simplex of $\calT'$.
    This implies $K=1$.
    So any boundary face can only be contained in one single $n$-simplex.

@@

    \item 
    So it remains to study the topology of vertex stars. Suppose that $S = \{z\}$ is a singleton simplex. 
    For $\delta > 0$ small enough, 
    \begin{align*}
        B_{\delta}(z) = \left\{ x \in \calM \suchthat d(x,z) \leq \delta \right\},
        \quad 
        S_{\delta}(z) = \left\{ x \in \calM \suchthat d(x,z) = \delta \right\},
    \end{align*}
    are compactly contained subsets of $|\patch_{\calT}(S)|$.
    For each simplex $K \in \patch_{\calT}(S)$ that contains the vertex $z$ and which has the opposite face $F_K$, 
    we easily find homeomorphisms $\phi_{K} : B_{\delta}(z) \cap K \rightarrow K$
    such that 
    $\phi_{K}(z) = z$, 
    such that 
    $\phi_{K}( K \cap S_{\delta}(z) ) = F_{K}$,
    such that 
    for all $K, K' \in \calT$ with $K' \subseteq K$ we have $\phi_{K|K'} = \phi_{K'}$,
    and such that 
    $\phi_{K}$ maps $[0,1]\cdot v$ linearly into $\bbR \cdot v$. 
    This is defines a homeomorphism $\phi : B_{\delta}(z) \rightarrow |\patch_{\calT}(S)|$.

@@

    If $z$ is a boundary vertex of $\calS'$,
    which is the case if and only if $S$ is a boundary simplex, 
    then the underlying space of $\carapace(S)$ is homeomorphic to $S_{\delta}(z) \cap |\patch_{\calT}(S)|$ along $\phi$.
    The vertex star around $z$ within the boundary triangulation $\partial\calT$ is homeomorphic to the ball of dimension $n-1$,
    and it is homeomorphic to $\overline{ B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)|$ along $\phi$. Notice that 
    \begin{align*}
        \partial\left( { B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)| \right)
        =
        S_{\delta}(z) \cap |\patch_{\calT}(S)|
        \cup 
        \overline{ B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)|
        .
    \end{align*}
    Provided that $\delta > 0$ is small enough, there exists a homeomorphism $\phi'$ from $\overline{B_{\delta}(z)} \cap |\calT|$
    onto the closed unit ball. Necessarily, this homeomorphism maps the boundary of $\overline{B_{\delta}(z)} \cap |\calT|$, which contains $z$, onto the unit sphere. Since the image of $\overline{ B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)|$ along $\phi'$ is still homeomorphic to a disk of dimension $n-1$, so must be the image of $S_{\delta}(z) \cap |\patch_{\calT}(S)|$ under $\phi'$.
    But that just means that $\carapace(S)$ triangulates a topological ball of dimension $n-1$.

>>>>

\begin{lemma}\label{lemma:boundarysimplices}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$. 
    \begin{itemize}
        \item Any face $F \in \Faces(\calT)$ is not contained in the boundary if and only if it is contained in exactly two $n$-simplices.
        \item Any face $F \in \Faces(\calT)$ is contained in the boundary if and only if it is contained in exactly one $n$-simplex.
        \item The simplices contained in the boundary constitute a triangulation of the boundary.
        % \item $\calT$ is non-branching, that is, every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$. 
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove these statements in several steps.
    \begin{enumerate}
    \item 
    We $\mathring \calM := \calM \setminus \partial \calM$ denote the interior of the manifold. 
    We will use the following fact:\footnote{To see this, one easily constructs a continuous deformation of $\calM$ into itself to move any chosen point on $S$ to any other chosen point on $S$.} if $S \in \calT$ has an inner point that lies on $\partial \calM$, then all inner points of $S$ are on $\partial \calM$. 
    Since the boundary $\partial \calM$ is closed, every $S \in \calT$ is either a subset of the boundary or all its inner points lie in the interior $\mathring \calM$ of the manifold.

    \item 
    We recall an auxiliary result.
    Suppose that $Y$ is a topological space homeomorphic to a sphere of dimension $m$ and that $X \subseteq Y$ is homeomorphic to a sphere of dimension $m-1$, where $m \geq 1$. 
    As a consequence of the Jordan-Brouwer separation theorem~\cite[Corollary IV.5.24]{mayer1989algebraische}~\cite[Corollary VIII.6.4]{massey1981algebraic}, 
    we know that $Y \setminus X$ has got two connected components.

    % \item 
    % We prove an auxiliary result. 
    % Suppose that $Y$ is a topological space homeomorphic to a sphere of dimension $m$ and that $X \subseteq Y$ is homeomorphic to a sphere of dimension $m-1$, where $m \geq 1$. 
    % The embedding of $X$ into $Y$ cannot be onto~\cite[Theorem~2.55]{lee2011topological}, and so there exists $y \in Y \setminus X$. 
    % Using the homeomorphism from $Y$ into the $m$-sphere, which maps $y$ to the northpole without loss of generality, and a stereographic projection, 
    % we see that $Y \setminus \{y\}$ is homeomorphic to $R^{m}$.
    % As a consequence of the Jordan-Brouwer separation theorem~\cite[Theorem IV.5.23]{mayer1989algebraische}~\cite[Corollary IV.5.24]{mayer1989algebraische}~\cite[Corollary VIII.6.4]{massey1981algebraic}, $Y \setminus (X \cup \{y\})$ has got two connected components,
    % and thus $Y \setminus X$ has got two connected components.

@@

    Consider the case that $z_F$ lies on the boundary of $\calM$ and suppose that $F$ is contained in $K$ distinct $n$-simplices of $\calT$.
    By adding at least one dimension, we can double\footnote{The reader is referred to Lee's textbook~\cite{lee2012smooth} for more background and the technicalities.} the manifold $\calM$ along the boundary and obtain the doubled manifold $\calM'$. 
    Similarly, we can construct a doubling of the triangulation $\calT'$ such that $F$ is contained in exactly $2K$ distinct $n$-simplices of $\calT'$. 
    We know that $\calM'$ is a manifold without boundary, and hence $F$ is an interior simplex of $\calT'$.
    This implies $K=1$.
    So any boundary face can only be contained in one single $n$-simplex.

@@

    \item 
    So it remains to study the topology of vertex stars. Suppose that $S = \{z\}$ is a singleton simplex. 
    For $\delta > 0$ small enough, 
    \begin{align*}
        B_{\delta}(z) = \left\{ x \in \calM \suchthat d(x,z) \leq \delta \right\},
        \quad 
        S_{\delta}(z) = \left\{ x \in \calM \suchthat d(x,z) = \delta \right\},
    \end{align*}
    are compactly contained subsets of $|\patch_{\calT}(S)|$.
    For each simplex $K \in \patch_{\calT}(S)$ that contains the vertex $z$ and which has the opposite face $F_K$, 
    we easily find homeomorphisms $\phi_{K} : B_{\delta}(z) \cap K \rightarrow K$
    such that 
    $\phi_{K}(z) = z$, 
    such that 
    $\phi_{K}( K \cap S_{\delta}(z) ) = F_{K}$,
    such that 
    for all $K, K' \in \calT$ with $K' \subseteq K$ we have $\phi_{K|K'} = \phi_{K'}$,
    and such that 
    $\phi_{K}$ maps $[0,1]\cdot v$ linearly into $\bbR \cdot v$. 
    This is defines a homeomorphism $\phi : B_{\delta}(z) \rightarrow |\patch_{\calT}(S)|$. 
    Moreover, one easily verifies that $\phi$ is bi-Lipschitz.
    For $\delta > 0$ small enough, $B_{\delta}(z)$ is homeomorphic to the unit ball and that homeomorphism maps $S_{\delta}(z)$ onto the unit sphere.

@@

    Now suppose that $z$ is a boundary vertex of $\calS'$,
    which is the case if and only if $S$ is a boundary simplex. 
    Write 
    \[
        B_{H} = \left\{ x \in B_1(0) \suchthat* x_1 \leq 0 \right\},
        \quad 
        S_{H} = \left\{ x \in \partial B_1(0) \suchthat* x_1 \leq 0 \right\}.
        \quad 
        D_{H} = \left\{ x \in \partial B_1(0) \suchthat* x_1 = 0 \right\}.
    \]
    Provided that $\delta > 0$ is small enough, there exists a homeomorphism $\phi'$ that maps $\overline{B_{\delta}(z)} \cap \calM$ onto $B_1(0)$, which maps $\overline{B_{\delta}(z)} \cap \partial\calM$ onto $D_H$,
    and which maps $z$ to the origin. 
    %
    It follows that $\phi'$ maps the closure of $\partial{B_{\delta}(z)} \setminus \partial\calM$ onto $S_H$.
    %
    Assuming that $\delta$ is small enough once more, 
    we now verify that $\phi^{-1}$ maps $|\carapace(S)|$ onto $S_{\delta}(z) \cap |\patch_{\calT}(S)|$.
    The latter equals the closure of $\partial{B_{\delta}(z)} \setminus \partial\calM$.
    It follows that $\carapace(S)$ triangulates a topological ball of dimension $n-1$.

    % We let $\calM'$ be the doubling of the manifold and let $\calT'$ be the doubling of the present triangulation;
    % we refer to Lee's textbook~\cite{lee2011topological} for details. 
    % The vertex star around $z$ within the boundary triangulation $\partial\calT$ is homeomorphic to the ball of dimension $n-1$,
    % and it is homeomorphic to $\overline{ B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)|$ along $\phi$. 
    % The underlying space of $\carapace(S)$ is homeomorphic to $S_{\delta}(z) \cap |\patch_{\calT}(S)|$ along $\phi$.
    % The vertex star around $z$ within the boundary triangulation $\partial\calT$ is homeomorphic to the ball of dimension $n-1$,
    % and it is homeomorphic to $\overline{ B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)|$ along $\phi$. 
    % Notice that 
    % \begin{align*}
    %     \partial\left( { B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)| \right)
    %     =
    %     S_{\delta}(z) \cap |\patch_{\calT}(S)|
    %     \cup 
    %     \overline{ B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)|
    %     .
    % \end{align*}
    % Necessarily, this homeomorphism maps the boundary of $\overline{B_{\delta}(z)} \cap |\calT|$, which contains $z$, onto the unit sphere. Since the image of $\overline{ B_{\delta}(z) } \cap |\partial\patch_{\calT}(S)|$ along $\phi'$ is still homeomorphic to a disk of dimension $n-1$, so must be the image of $S_{\delta}(z) \cap |\patch_{\calT}(S)|$ under $\phi'$.
    % But that just means that $\carapace(S)$ triangulates a topological ball of dimension $n-1$.
